---
title: "PreprocessingTrainingDatasetforTRN"
author: "Alison Paquette"
date: "9/23/2020"
date Written: "September 2019"
date Code Streamlined: "6/20/2020"
output: html_document
---

The goal of this code is to take the Enhancer and No Enhaner base models that Paul generated in September 2019 and get them to a state where we can run pearson correlations within our training datset.  The actual correlations are run in parellel using a custom R script 


#### B. Load Input Data ####
```{r}
load("/local/users/apaquett/PlacentalTRN/IntermediateData/TrainingDataset_6192020.RData")
load("/local/users/apaquett/PlacentalTRN/PreprocessedData/EnhancerBaseTRN.RData")  #Note: Ignore all of the data in this model. This is just to extract the potential binding sites out of what pauls code is set up to run.
colnames(EnhancerFullModel)[1]<-"tf"

load("/local/users/apaquett/PlacentalTRN/PreprocessedData/NoEnhancerBaseTRN.RData")  #Note: Ignore all of the data in this model. This is just to extract the potential binding sites out of what pauls code is set up to run.
colnames(NoEnhancerFullModel)[1]<-"tf"

##This is fromt he Human TF DB (Lambert et al, 2018
DBTFs<-read.csv("~/RandomImpFiles/HumanTFs.csv")
rownames(DBTFs)<-DBTFs$HGNC.symbol
DBTFs<-subset(DBTFs,Is.TF.=="Yes")
DBTFs<-as.character(DBTFs[,2])
```


#### C. Make No Enhancer Model to Get Correlations ####

```{r}
TGs<-as.data.frame(table(NoEnhancerFullModel$targetGene))
colnames(TGs)<-c("Target","NTFs")
rownames(TGs)<-TGs$Target

TFsList<- list()
for(i in 1: nrow(TGs)){
  TFsList[[i]] <- NoEnhancerFullModel[NoEnhancerFullModel$targetGene==rownames(TGs)[i],]$tf
}
names(TFsList) <-TGs$Target

TFs<-as.data.frame(table(NoEnhancerFullModel$tf))
colnames(TFs)<-c("TF","N_Targets")
rownames(TFs)<-TFs$TF


TGsList<- list()
for(i in 1: nrow(TFs)){
  TGsList[[i]] <- NoEnhancerFullModel[NoEnhancerFullModel$tf==rownames(TFs)[i],]$targetGene
}
names(TGsList) <- TFs$TF

#Remove whatever discrepencies there were with the dummy dataset and this one
MatchedTGs<-intersect(rownames(Exprs_Train),names(TFsList))
TGs<-TGs[MatchedTGs,]
TFsList<-TFsList[MatchedTGs]
names(TGs)

#We are including ONLY TFs defined by the Human TF Database (Lambert et al, 2018)
MatchedTFs<-intersect(DBTFs,names(TGsList))
TFs<-TFs[MatchedTFs,]
TGsList<-TGsList[MatchedTFs]
#save(TGsList,TFsList,TFs,TGs,file="~/PlacentalTRN/PreprocessedData/NoEnhancerProcessedModel.RData")

```


################################################################################################
##  Run on Server: Rscript ServerCode_PearsonNoEnhancerTrainingData.R ##########################
################################################################################################


#### C. Make Enhancer Model to Get Correlations ####

```{r}
TGs<-as.data.frame(table(EnhancerFullModel$targetGene))
colnames(TGs)<-c("Target","NTFs")
rownames(TGs)<-TGs$Target
#TGs<-subset(TRNFiltTest_TGs,NTFs>3)#For lasso regression, we need > 2 in matrix

#Making these lists takes X (Started 3:40 PM; Probably took ~ 30 Minutes)
TFsList<- list()
for(i in 1: nrow(TGs)){
  TFsList[[i]] <- EnhancerFullModel[EnhancerFullModel$targetGene==rownames(TGs)[i],]$tf
}
names(TFsList) <-TGs$Target


TFs<-as.data.frame(table(EnhancerFullModel$tf))
colnames(TFs)<-c("TF","N_Targets")
rownames(TFs)<-TFs$TF


TGsList<- list()
for(i in 1: nrow(TFs)){
  TGsList[[i]] <- EnhancerFullModel[EnhancerFullModel$tf==rownames(TFs)[i],]$targetGene
}
names(TGsList) <- TFs$TF

#Remove whatever discrepencies there were with the dummy dataset and this one
MatchedTGs<-intersect(rownames(Exprs_Train),names(TFsList))
TGs<-TGs[MatchedTGs,]
TFsList<-TFsList[MatchedTGs]
names(TGs)

#We are including ONLY TFs defined by the Human TF Database (Lambert et al, 2018)
MatchedTFs<-intersect(DBTFs,names(TGsList))
TFs<-TFs[MatchedTFs,]
TGsList<-TGsList[MatchedTFs]
#save(TGsList,TFsList,TFs,TGs,file="~/PlacentalTRN/PreprocessedData/EnhancerProcessedModel.RData")
```



################################################################################################
##  Run on Server: Rscript ServerCode_PearsonEnhancerTrainingData.R ##########################
################################################################################################
