---
title: "PretermBirthAnalysis"
author: "Alison Paquette"
date: "9/30/2020"
output: html_document
---

# Load Relevant Data
-load the TRN as it will be presented in the R shiny app so that we can ensure
```{r}
TRNModel<-read.csv("~/PlacentalTRN/FinalResults/TRNForRShinyApp.csv")
TRNModel<-TRNModel[,-1]
TRNModel$targetGene<-as.character(TRNModel$targetGene)
TRNModel$TF<-as.character(TRNModel$TF)

#Load results of miroarray analysis published in BOR Paper
load("~/PlacentalTRN/PreprocessedData/AggregatedMicroarraySig_BOR2017.RData")
Microarray_sig$hgnc_symbol<-as.character(Microarray_sig$hgnc_symbol)
rownames(Microarray_sig)<-as.character(Microarray_sig$hgnc_symbol)

load("~/PlacentalTRN/FinalResults/Node&Edge.RData")
```


# Enrichment Test Function####

Perform Fishers exact test to see if Gene list is enriched for specifif TFs.
Reference: https://www.pathwaycommons.org/guide/primers/statistics/fishers_exact_test/
```{r}
EnrichFisher<-function(NSigGenes,NTotalGenes,SigInGroup,GroupSize){

  a=SigInGroup ## Equivalent to X; Table 3. In the TFgroup and DE
  b= GroupSize-SigInGroup #. Equivalent to m-X: In the TFGroup but NOT DE
  c=NSigGenes-SigInGroup  # Equivalent to K-X; DE but not in the group
  n=NTotalGenes-GroupSize #n in contigencyTable: not in group
  d=n-c
  mat<-(matrix(c(a,c,b,d),nrow=2))
  mat
  sum(mat)==NTotalGenes
  fisher.test(mat,alternative="greater")$p.value #default is 2 sided. We need this to be greater rather than less to say that these genes are enriched
}
```

# Generate TF Enrichment Table
-This is basically what the RShiny function is doing
```{r}



TFTable<-as.data.frame(table(TRNModel$TF))
colnames(TFTable)<-c("TF","NTargets")
rownames(TFTable)<-TFTable$TF
TFTable$TF<-as.character(TFTable$TF)
TFTable$SigGenes<-NA

  for(i in 1:dim(TFTable)[1])
  {
    x<-subset(TRNModel,TF==as.character(TFTable$TF[i])) # Pull out the TF of interest
    dim(x)[1]==TFTable$NTargets[i] #This should be equal
    TFTable$SigGenes[i]<-length(intersect(as.character(x$targetGene),Microarray_sig$hgnc_symbol)) #How many in the input gene list match this TF?
  }



  TFTable$FisherTest<-NA
  # Perform Enrichment Test using Fishers Exact Function
  for(i in 1:length(TFTable$TF)){
    TFTable$FisherTest[i]<-EnrichFisher(NSigGenes=as.numeric(length(Microarray_sig$hgnc_symbol)),NTotalGenes=7712,SigInGroup =TFTable$SigGenes[i],GroupSize=TFTable$NTargets[i])
  }

AllTFTable<-TFTable
# Filter By Preset Thresholds

MinSigGenes=3
#ModuleMax

#TFTable<-subset(TFTable,NTargets>=ModuleMin)
#TFTable<-subset(TFTable,NTargets<ModuleMax)
TFTable<-subset(TFTable,SigGenes>=MinSigGenes)
TFTable$FisherAdjustP<-p.adjust(TFTable$FisherTest,method="BH",n=dim(TFTable)[1])
```

#Pull out Enriched TFs and get calculations for paper
```{r}
EnrichedTFs<-subset(TFTable,FisherTest<0.1)$TF

#1. How many total Genes in List?
GenesIncluded<-NULL
for(i in 1:length(EnrichedTFs)){
  TargetGenes<-as.character(subset(TRNModel,TF==EnrichedTFs[i])$targetGene)
  TargetGenes<-intersect(TargetGenes,Microarray_sig$hgnc_symbol)
  GenesIncluded<-union(TargetGenes,GenesIncluded)
}


length(GenesIncluded)/length(Microarray_sig$hgnc_symbol)

Table4<-TFTable[EnrichedTFs,]

Table4$Genes<-NA

for(i in 1:length(EnrichedTFs)){
    TargetGenes<-as.character(subset(TRNModel,TF==EnrichedTFs[i])$targetGene)
    TargetGenes<-intersect(TargetGenes,Microarray_sig$hgnc_symbol)
    Table4$Genes[i]<-paste(TargetGenes,collapse=" ")
}

write.table(Table4,file="~/PlacentalTRN/FinalResults/Table4EnrichedTFs.csv")

```

# Which Preterm birth genes were ALSO TFs?
```{r}
rownames(AllTFTable)<-AllTFTable$TF
PTBTFs<-AllTFTable[intersect(Microarray_sig$hgnc_symbol,AllTFTable$TF),]

PTBTFs<-merge(PTBTFs,Microarray_sig,by='row.names',all.x=T,all.y=F)

write.table(PTBTFs,file="~/PlacentalTRN/FinalResults/PTB_TFs.csv")

```

# Make Culled network for Cytoscape
```{r}
### Fix nodes and Edges and Export to Cytoscape
#Pull out Enriched TFs
Edge_MA<-t(as.data.frame(rep(NA,7),ncols=7,nrows=1))
colnames(Edge_MA)<-colnames(Edge)
for(i in 1:length(EnrichedTFs)){
  temp1<- subset(Edge, TF == EnrichedTFs[i])
  rownames(temp1)<-temp1$targetGene
  temp2<-temp1[intersect(Microarray_sig$hgnc_symbol,temp1$targetGene),]
  rownames(temp2)<-paste(temp2$TF,temp2$targetGene)
  Edge_MA<-rbind(Edge_MA,temp2)
}

##Pull out connections Between TFs that might match
for(i in 1:length(EnrichedTFs)){
  temp1<- subset(Edge, TF == EnrichedTFs[i])
  rownames(temp1)<-temp1$targetGene
  temp1$targetGene<-as.character(temp1$targetGene)
  temp2<-temp1[intersect(EnrichedTFs,temp1$targetGene),]
  rownames(temp2)<-paste(temp2$TF,temp2$targetGene)
  Edge_MA<-rbind(Edge_MA,temp2)
}

Edge_MA<-na.omit(Edge_MA)

Node_MA<-Node[c(unique(Edge_MA$TF),unique(Edge_MA$targetGene)),]

Node_MA$PTB<-rep("NO",dim(Node_MA)[1])
Node_MA$PTB[match(Microarray_sig$hgnc_symbol,Node_MA$Gene)]<-"YES"
table(Node_MA$PTB)

Node_MA<-merge(Node_MA,Microarray_sig,by.x="Gene",by.y="hgnc_symbol",all.x=T,all.y=F)


Node_MA<-Node_MA[,-c(6,7,9,11)]


Node_MA$LFCDir<-Node_MA$logFC
Node_MA$LFCDir[Node_MA$LFCDir>0] <- "POS"
Node_MA$LFCDir[Node_MA$LFCDir<=0] <- "NEG"
table(Node_MA$LFCDir)


Node_MA$Enriched<-rep("NO",dim(Node_MA)[1])
Node_MA$Enriched[match(EnrichedTFs,Node_MA$Gene)]<-"YES"
table(Node_MA$Enriched)


write.csv(Node_MA,file="~/PlacentalTRN/FinalResults/PTBMicroarray_Node_9222020.csv")
write.csv(Edge_MA,file="~/PlacentalTRN/FinalResults/PTBMicroarray_Edge_9222020.csv")

```

