---
title: "TrainingData-Model Parameterization"
author: "Alison Paquette"
date Written: "September 2019"
date Modified: "6/20/2020"
date RMarkdown Created: "9/23/2020"
output: html_document
---

#### B. Load Data ####
```{r}
library(foreach)
library(doParallel)
load("/local/users/apaquett/PlacentalTRN/IntermediateData/NoEnhancerCorrelationsTrainingData.RData")
load("/local/users/apaquett/PlacentalTRN/IntermediateData/EnhancerCorrelationsTrainingData.RData")
load("/local/users/apaquett/PlacentalTRN/IntermediateData/TrainingDataset_6192020.RData")
```

Set up a blank database to insert the model parameters into
```{r}

ModelDimensions = data.frame(matrix(vector(), 9, 5, dimnames=list(c(), c("NInteractions","TFs","TGs","MedianModuleSize","BiggestModule"))),stringsAsFactors=F)
rownames(ModelDimensions)<-c("Model1","Model2","Model3","Model4","Model5","Model6","Model7","Model8","Model9")
```


#### Model 2 
Use Median Correlation Cutoff; No Enhancers
```{r}
Mean<-median((abs(NoEnhancer_Cor$Cor)))
Cutoffs<-c((-Mean),Mean)
Model2<-subset(NoEnhancer_Cor,Cor<Cutoffs[1]|Cor>Cutoffs[2])
Model2<-subset(Model2,Rank<16)
hist(NoEnhancer_Cor$Cor,breaks=100)
hist(Model2$Cor,breaks=100)

TGs<-as.data.frame(table(Model2$Target_Gene)) 
colnames(TGs)<-c("Target","NTFs")
rownames(TGs)<-TGs$Target
TGs<-subset(TGs,NTFs>3)#For lasso regression, we need > 2 in matrix

TFs<-as.data.frame(table(Model2$TF))
colnames(TFs)<-c("TF","N_Targets")
rownames(TFs)<-TFs$TF


TFsList<- list()
for(i in 1: nrow(TGs)){
  TFsList[[i]] <- Model2[Model2$Target_Gene==rownames(TGs)[i],]$TF
}
names(TFsList) <-TGs$Target

ModelDimensions[2,]<-c(dim(Model2)[1], #N Interactoins
                       length(unique(as.character(Model2$TF))), #Number of TFs
                       dim(TGs)[1], #Number of TGs
                       median(TFs$N_Targets), #Median Number of TGs
                       max(TFs$N_Targets)) # Maximum Module Size

#save(Model2,TFsList,TGs,file="~/PlacentalTRN/IntermediateData/TrainingDataModel2_Full.RData")

```

#### Model 1 
** BASE MODEL **
Use median correlation cutoff, and Enhancers.
```{r}
Mean<-median((abs(Enhancer_Cor$Cor)))
Cutoffs<-c((-Mean),Mean)
Model1<-subset(Enhancer_Cor,Cor<Cutoffs[1]|Cor>Cutoffs[2])
Model1<-subset(Model1,Rank<16)
hist(Enhancer_Cor$Cor,breaks=100)
hist(Model1$Cor,breaks=100)

TGs<-as.data.frame(table(Model1$Target_Gene))    
colnames(TGs)<-c("Target","NTFs")
rownames(TGs)<-TGs$Target
TGs<-subset(TGs,NTFs>3)#For lasso regression, we need > 2 in matrix

TFsList<- list()
for(i in 1: nrow(TGs)){
  TFsList[[i]] <- Model1[Model1$Target_Gene==rownames(TGs)[i],]$TF
}
names(TFsList) <-TGs$Target

ModelDimensions[1,]<-c(dim(Model1)[1], #N Interactoins
                       length(unique(as.character(Model1$TF))), #Number of TFs
                       dim(TGs)[1], #Number of TGs
                       median(TFs$N_Targets), #Median Number of TGs
                       max(TFs$N_Targets)) # Maximum Module Size

save(Model1,TFsList,TGs,file="~/PlacentalTRN/IntermediateData/TrainingDataModel1_Full.RData")


```

#### Model 3
Use correlation cutoff of 0.25, 15 Tfs max, and enhancers
```{r}
Cutoffs<-c((-0.25),0.25)

Model3<-subset(Enhancer_Cor,Cor<Cutoffs[1]|Cor>Cutoffs[2])
Model3<-subset(Model3,Rank<16)
hist(Enhancer_Cor$Cor,breaks=100)
hist(Model3$Cor,breaks=100)



TGs<-as.data.frame(table(Model3$Target_Gene))
colnames(TGs)<-c("Target","NTFs")
rownames(TGs)<-TGs$Target
TGs<-subset(TGs,NTFs>3)#For lasso regression, we need > 2 in matrix

TFs<-as.data.frame(table(Model3$TF))
colnames(TFs)<-c("TF","N_Targets")
rownames(TFs)<-TFs$TF

TFsList<- list()
for(i in 1: nrow(TGs)){
  TFsList[[i]] <- Model3[Model3$Target_Gene==rownames(TGs)[i],]$TF
}
names(TFsList) <-TGs$Target

ModelDimensions[3,]<-c(dim(Model3)[1], #N Interactoins
                       length(unique(as.character(Model3$TF))), #Number of TFs
                       dim(TGs)[1], #Number of TGs
                       median(TFs$N_Targets), #Median Number of TGs
                       max(TFs$N_Targets)) # Maximum Module Size

#save(Model3,TFsList,TGs,file="~/PlacentalTRN/IntermediateData/TrainingDataModel3_Full.RData")
```



#######___Model 4 ############
Use a correlation cutoff of 0.5, only 15 TFs and Enhancers
```{r}
Cutoffs<-c((-0.5),0.5)
Model4<-subset(Enhancer_Cor,Cor<Cutoffs[1]|Cor>Cutoffs[2])
Model4<-subset(Model4,Rank<16)

hist(Enhancer_Cor$Cor,breaks=100)
hist(Model4$Cor,breaks=100)

TGs<-as.data.frame(table(Model4$Target_Gene))
colnames(TGs)<-c("Target","NTFs")
rownames(TGs)<-TGs$Target
TGs<-subset(TGs,NTFs>3)#For lasso regression, we need > 2 in matrix

TFs<-as.data.frame(table(Model4$TF))
colnames(TFs)<-c("TF","N_Targets")
rownames(TFs)<-TFs$TF


TFsList<- list()
for(i in 1: nrow(TGs)){
  TFsList[[i]] <- Model4[Model4$Target_Gene==rownames(TGs)[i],]$TF
}
names(TFsList) <-TGs$Target

ModelDimensions[4,]<-c(dim(Model4)[1], #N Interactoins
                       length(unique(as.character(Model4$TF))), #Number of TFs
                       dim(TGs)[1], #Number of TGs
                       median(TFs$N_Targets), #Median Number of TGs
                       max(TFs$N_Targets)) # Maximum Module Size

#save(Model4,TFsList,TGs,file="~/PlacentalTRN/IntermediateData/TrainingDataModel4_Full.RData")


```

#### Model 
Use only top 10 TFs instead of 15. still only > median correlation cutoff and enhnacers

```{r}
ean<-mean((abs(Enhancer_Cor$Cor)))
Cutoffs<-c((-Mean),Mean)
Model5<-subset(Enhancer_Cor,Cor<Cutoffs[1]|Cor>Cutoffs[2])
Model5<-subset(Model5,Rank<11)

TFs<-as.data.frame(table(Model5$TF))
colnames(TFs)<-c("TF","N_Targets")
rownames(TFs)<-TFs$TF


hist(Enhancer_Cor$Cor,breaks=100)
hist(Model5$Cor,breaks=100)

TGs<-as.data.frame(table(Model5$Target_Gene))     
colnames(TGs)<-c("Target","NTFs")
rownames(TGs)<-TGs$Target
TGs<-subset(TGs,NTFs>3)#For lasso regression, we need > 2 in matrix

TFsList<- list()
for(i in 1: nrow(TGs)){
  TFsList[[i]] <- Model5[Model5$Target_Gene==rownames(TGs)[i],]$TF
}
names(TFsList) <-TGs$Target

ModelDimensions[5,]<-c(dim(Model5)[1], #N Interactoins
                       length(unique(as.character(Model5$TF))), #Number of TFs
                       dim(TGs)[1], #Number of TGs
                       median(TFs$N_Targets), #Median Number of TGs
                       max(TFs$N_Targets)) # Maximum Module Size

#save(Model5,TFsList,TGs,file="~/PlacentalTRN/IntermediateData/TrainingDataModel5_Full.RData")

```

#### Model 6
```{r}
ean<-median((abs(Enhancer_Cor$Cor)))
Cutoffs<-c((-Mean),Mean)
Model6<-subset(Enhancer_Cor,Cor<Cutoffs[1]|Cor>Cutoffs[2])
Model6<-subset(Model6,Rank<11)

TFs<-as.data.frame(table(Model6$TF))
colnames(TFs)<-c("TF","N_Targets")
rownames(TFs)<-TFs$TF


hist(Enhancer_Cor$Cor,breaks=100)
hist(Model6$Cor,breaks=100)

TGs<-as.data.frame(table(Model6$Target_Gene))
colnames(TGs)<-c("Target","NTFs")
rownames(TGs)<-TGs$Target
TGs<-subset(TGs,NTFs>3)#For lasso regression, we need > 2 in matrix

TFsList<- list()
for(i in 1: nrow(TGs)){
  TFsList[[i]] <- Model6[Model6$Target_Gene==rownames(TGs)[i],]$TF
}
names(TFsList) <-TGs$Target

ModelDimensions[6,]<-c(dim(Model6)[1], #N Interactoins
                       length(unique(as.character(Model6$TF))), #Number of TFs
                       dim(TGs)[1], #Number of TGs
                       median(TFs$N_Targets), #Median Number of TGs
                       max(TFs$N_Targets)) # Maximum Module Size

#save(Model6,TFsList,TGs,file="~/PlacentalTRN/IntermediateData/TrainingDataModel6_Full.RData")


```



#######BAD MODELS  ############

We also included some false models in our database which had no binding sites for TFs for each gene and/or poor correlations.  


1. make Fake & Real Data Matrix: Note We wanna be using the same thing for all models so it s easier to run on server 
```{r}
TGs<-as.data.frame(table(Enhancer_Cor$Target_Gene))
colnames(TGs)<-c("Target","NTFs")
rownames(TGs)<-TGs$Target
TGs<-subset(TGs,NTFs>3)#For lasso regression, we need > 2 in matrix

TFs<-as.data.frame(table(Enhancer_Cor$TF))
colnames(TFs)<-c("TF","N_Targets")
TFs$TF<-as.character(TFs$TF)


TFsList_Real<- list()
TFsList_Fake<- list()
for(i in 1: nrow(TGs)){
  TFsList_Real[[i]]<-Enhancer_Cor[Enhancer_Cor$Target_Gene==rownames(TGs)[i],]$TF
  TFsList_Fake[[i]]<-setdiff(TFs$TF,TFsList_Real[[i]])
}
names(TFsList_Real) <- as.character(TGs$Target)
names(TFsList_Fake) <- as.character(TGs$Target)


#save(TFsList_Fake,TFsList_Real,file="~/PlacentalTRN/IntermediateData/EnhancerModels_FakeandRealMatchedTFs.RData")

Corr_Results<- function(Exprs_Train,TargetGene,TFs_Fake){
  Corr<-data.frame(matrix(NA,nrow=length(TFs_Fake),ncol=1))
  for (i in 1:length(TFs_Fake)){
    Corr[i,1]<-cor.test(as.numeric(Exprs_Train[TargetGene,]),as.numeric(Exprs_Train[TFs_Fake[i],]))$estimate
  }
  Corr$Target_Gene<-TargetGene
  Corr$TF<-TFs_Fake
  Corr
}


NCores=4
cl <- makeCluster(NCores) #not to overload your computer
registerDoParallel(cl)

# Parellelized Code
Fake_Cor<- foreach(j=1:length(rownames(TGs)), .combine=rbind,.errorhandling='remove') %dopar% {
  tempMatrix = Corr_Results(Exprs_Train,names(TFsList_Fake)[j],TFsList_Fake[[j]]) #calling a function
  tempMatrix #Equivalent to finalMatrix = cbind(finalMatrix, tempMatrix)
}

colnames(Fake_Cor)[1]<-"Cor"


Fake_Cor<-subset(Fake_Cor,Cor<0.99999) #This gets rid of auto correlatoins
####
#save(Fake_Cor,file="~/PlacentalTRN/IntermediateData/FakeCorrelations_TrainingData.RData")

```

#  We need to plot this data to make sure the cutoffs are correct

```{r}
p3 <- hist(abs(Enhancer_Cor$Cor),breaks=40)                   # centered at 4
p4 <- hist(abs(Fake_Cor$Cor), breaks=40)                  # centered at 6
plot( p3, col="orange", xlim=c(0,1),main="Absolute Value: Pearson Correlation Coefficents")  # first histogram
plot( p4, col="blue", xlim=c(0,1),add=T)  # second
mean_Real<-mean(abs(Enhancer_Cor$Cor))
mean_Fake<-mean(abs(Fake_Cor$Cor))
abline(v=mean_Real,col="orange",lwd=2,lty=2)
abline(v=mean_Fake,col="blue",lwd=2,lty=2)

t.test(abs(Enhancer_Cor$Cor),abs(Fake_Cor$Cor))
```

#### Bad Model 7: Real Binding, Bad correlations #########

```{r}
Mean<-median((abs(Enhancer_Cor$Cor)))
Cutoffs<-c((-Mean),Mean)
Real_BadCors<-subset(Enhancer_Cor,Cor<Cutoffs[2])
Real_BadCors<-subset(Real_BadCors,Cor>Cutoffs[1])


Mean<-median((abs(Enhancer_Cor$Cor)))
Cutoffs<-c((-Mean),Mean)
Real_BadCors<-subset(Enhancer_Cor,Cor<Cutoffs[2])
Real_BadCors<-subset(Real_BadCors,Cor>Cutoffs[1])


TGs<-as.data.frame(table(Enhancer_Cor$Target_Gene))    
colnames(TGs)<-c("Target","NTFs")
rownames(TGs)<-TGs$Target
#TGs<-subset(TGs,NTFs>3)#For lasso regression, we need > 2 in matrix

FakeOptions<-unique(Real_BadCors$Target_Gene)
colnames(Real_BadCors)[1]<-"tf"
NoMatches<-setdiff(as.character(TGs$Target),FakeOptions)

TGs<-TGs[intersect(as.character(TGs$Target),FakeOptions),]


BAD_TFsList<- list() #Randomly Sample TFs with the same number as the target genes
for(i in 1: nrow(TGs)){
  TG<-as.character(TGs$Target[i])
  Options<-Real_BadCors[Real_BadCors$Target_Gene==as.character(TGs[i,]$Target),]
  
  if(length(Options$TF)<=(TGs[i,]$NTFs)){ #If there are less options then the TFs in the original model...
    BAD_TFsList[[i]] <- Options$TF
  } else {
    
    BAD_TFsList[[i]] <- sample(Options$TF,TGs[i,]$NTFs)
  }
  if(length(BAD_TFsList[[i]])<3){
    BAD_TFsList[[i]]=NULL
  }
}
names(BAD_TFsList) <- as.character(TGs$Target)



##Randomly sample and make sure no overlaps btween bad and good list. This should fail
#for(i in 1:5){
#any<-NA
#check<-sample(names(TFsList),5)
#any[i]<-intersect(BAD_TFsList[check[i]],TFsList[check[i]])
#any
#}

TFsList<-plyr::compact(BAD_TFsList)

##Note: These all need to be saaved with these names for the downsteram code to work. You needt o chagne the file name itself, bu tnot these names. NEVER CHANGE THESE NAMES

#save(TFsList,TGs,file="~/PlacentalTRN/IntermediateData/TrainingDataModel7_REALBADCORS.RData")


```

# Model 8
Fake Binding Sites but Good correlations

```{r}
#########################################################
#### BadModel 8 : Fake Binding, Good correlations #########
#########################################################
rm(list = ls())
load("/local/users/apaquett/PlacentalTRN/IntermediateData/EnhancerModels_FakeandRealMatchedTFs.RData")
load("/local/users/apaquett/PlacentalTRN/IntermediateData/FakeCorrelations_TrainingData.RData")
load("/local/users/apaquett/PlacentalTRN/IntermediateData/TrainingDataModel1_Full.RData")
load("/local/users/apaquett/PlacentalTRN/IntermediateData/EnhancerCorrelationsTrainingData.RData")




Median<-median((abs(Enhancer_Cor$Cor)))
Cutoffs<-c((-Median),Median)
Fake_GoodCors<-subset(Fake_Cor,Cor<Cutoffs[1]|Cor>Cutoffs[2])

Fake_BadCors<-subset(Fake_Cor,Cor<Cutoffs[2])#| Cor>Cutoff[1])
Fake_BadCors<-subset(Fake_BadCors,Cor>Cutoffs[1])


p3 <- hist(Fake_GoodCors$Cor,breaks=40)                   # centered at 4
p4 <- hist(Fake_BadCors$Cor, breaks=10)                  # centered at 6
plot( p4, col="orange", xlim=c(-1,1))  # first histogram
plot( p3, col="blue", xlim=c(-1,1),add=T)  # second

#should all be "in bounds"
table(Fake_BadCors$Cor<Cutoffs[1])
table(Fake_BadCors$Cor>Cutoffs[2])

#Should have equivalent out of bounds in the true and false category
table(Fake_GoodCors$Cor<Cutoffs[1])
table(Fake_GoodCors$Cor>Cutoffs[2])
# First get st up with filtering Criteria

FakeOptions<-unique(Fake_GoodCors$Target_Gene)

NoMatches<-setdiff(as.character(TGs$Target),FakeOptions)

TGs<-TGs[intersect(TGs$Target,FakeOptions),]


## NOTE: THIS TAKES A WHILE IF THERE ARE LOTS OF OPTIONS
BAD_TFsList<- list()
for(i in 1: nrow(TGs)){
  TG<-as.character(TGs[i,])
  Options<-Fake_GoodCors[Fake_GoodCors$Target_Gene==as.character(TGs[i,]$Target),]
  
  if(length(Options$TF)<=(TGs[i,]$NTFs)){ #If there are less options then the TFs in the original model...
    BAD_TFsList[[i]] <- Options$TF #just deal with whatever is available
  } else {
    
    BAD_TFsList[[i]] <- sample(Options$TF,TGs[i,]$NTFs) #If not-> randomly select a copule
  }
  if(length(BAD_TFsList[[i]])<3){
    TFsList[[i]]=NULL
  }
}

names(BAD_TFsList) <- TGs$Target

#for(i in 1:5){
  #any<-NA
  #check<-sample(names(TFsList),5)
  #any[i]<-intersect(BAD_TFsList[check[i]],TFsList[check[i]])
  #any
#}

TFsList<-plyr::compact(BAD_TFsList)

#(TFsList,TGs,file="~/PlacentalTRN/IntermediateData/TrainingDataModel8_FAKEGOODCORS.RData")

```

# Model 9: no binding sites and bad correlations
```{r}
rm(list = ls())
load("/local/users/apaquett/PlacentalTRN/IntermediateData/EnhancerModels_FakeandRealMatchedTFs.RData")
load("/local/users/apaquett/PlacentalTRN/IntermediateData/FakeCorrelations_TrainingData.RData")
load("/local/users/apaquett/PlacentalTRN/IntermediateData/TrainingDataModel1_Full.RData")
load("/local/users/apaquett/PlacentalTRN/IntermediateData/EnhancerCorrelationsTrainingData.RData")


Median<-median((abs(Enhancer_Cor$Cor)))
Cutoffs<-c((-Median),Median)
Fake_GoodCors<-subset(Fake_Cor,Cor<Cutoffs[1]|Cor>Cutoffs[2])

Fake_BadCors<-subset(Fake_Cor,Cor<Cutoffs[2])#| Cor>Cutoff[1])
Fake_BadCors<-subset(Fake_BadCors,Cor>Cutoffs[1])


p3 <- hist(Fake_GoodCors$Cor,breaks=40)                   # centered at 4
p4 <- hist(Fake_BadCors$Cor, breaks=10)                  # centered at 6
plot( p4, col="orange", xlim=c(-1,1))  # first histogram
plot( p3, col="blue", xlim=c(-1,1),add=T)  # second

#should all be "in bounds"
table(Fake_BadCors$Cor<Cutoffs[1])
table(Fake_BadCors$Cor>Cutoffs[2])

#Should have equivalent out of bounds in the true and false category
table(Fake_GoodCors$Cor<Cutoffs[1])
table(Fake_GoodCors$Cor>Cutoffs[2])
# First get st up with filtering Criteria

FakeOptions<-unique(Fake_BadCors$Target_Gene)

NoMatches<-setdiff(as.character(TGs$Target),FakeOptions)

TGs<-TGs[intersect(TGs$Target,FakeOptions),]

#TRNFiltTest_TFs<-as.data.frame(table(TRNFiltTest$tf))
#colnames(TRNFiltTest_TFs)<-c("TF","N_Targets")
#rownames(TRNFiltTest_TFs)<-TRNFiltTest_TFs$TF

## NOTE: THIS TAKES A WHILE IF THERE ARE LOTS OF OPTIONS
BAD_TFsList<- list()
for(i in 1: nrow(TGs)){
  TG<-as.character(TGs[i,])
  Options<-Fake_BadCors[Fake_BadCors$Target_Gene==as.character(TGs[i,]$Target),]
  
  if(length(Options$TF)<=(TGs[i,]$NTFs)){ #If there are less options then the TFs in the original model...
    BAD_TFsList[[i]] <- Options$TF #just deal with whatever is available
  } else {
    
    BAD_TFsList[[i]] <- sample(Options$TF,TGs[i,]$NTFs) #If not-> randomly select a copule
  }
  if(length(BAD_TFsList[[i]])<3){
    TFsList[[i]]=NULL
  }
}

names(BAD_TFsList) <- TGs$Target

#for(i in 1:5){
 # any<-NA
#  check<-sample(names(TFsList),5)
 # any[i]<-intersect(BAD_TFsList[check[i]],TFsList[check[i]])
 # any
#}

TFsList<-plyr::compact(BAD_TFsList)

#save(TFsList,TGs,file="~/PlacentalTRN/IntermediateData/TrainingDataModel9_FAKEBADCORS.RData")
```
